---
title: "statistical Analysis"
author: "Jisha"
date: "March 12, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Introduction
  
  #American Dream Home
 
  For many Americans homeownership is part of achieving their personal 'American Dream'. Most Millennial renters aged 18-34 (89%) plan to buy a home one day - more than any other generation. But what does this dream home look like? And where is located? What amenities do people dream of most?
For most people, a house is the biggest purchase they will make in their lives, one they will pay off for years, even decades, to come. Based on the type of homes people prefer and their priorities in amenities available in those homes answer the questions of new home buyers, sellers, remodelers and builders to make better decisions.

#Data
The data set is Kaggle's ,'Iowa House Prices: Regression Techniques'- train dataset.Dataset contains 81 variables and 1461 observations.Of the 81 variables, 23 are nominal, 23 are ordinal, 14 are discrete, and 21 are continuous. The variables included are basic characteristics that anyone wanting to buy a house would be interested in.

```{r}
hd <-  read.csv("C:/springboard_capstone_project/final_project/home_poject_data/train.csv", header = TRUE)
View(hd)

```


summary(homedf)
##        Id           MSSubClass       MSZoning     LotFrontage    
##  Min.   :   1.0   Min.   : 20.0   C (all):  10   Min.   : 21.00  
##  1st Qu.: 365.8   1st Qu.: 20.0   FV     :  65   1st Qu.: 59.00  
##  Median : 730.5   Median : 50.0   RH     :  16   Median : 69.00  
##  Mean   : 730.5   Mean   : 56.9   RL     :1151   Mean   : 70.05  
##  3rd Qu.:1095.2   3rd Qu.: 70.0   RM     : 218   3rd Qu.: 80.00  
##  Max.   :1460.0   Max.   :190.0                  Max.   :313.00  
##                                                  NA's   :259     
##     LotArea        Street      Alley      LotShape  LandContour
##  Min.   :  1300   Grvl:   6   Grvl:  50   IR1:484   Bnk:  63   
##  1st Qu.:  7554   Pave:1454   Pave:  41   IR2: 41   HLS:  50   
##  Median :  9478               NA's:1369   IR3: 10   Low:  36   
##  Mean   : 10517                           Reg:925   Lvl:1311   
##  3rd Qu.: 11602                                                
##  Max.   :215245                                                
##                                                                
##   Utilities      LotConfig    LandSlope   Neighborhood   Condition1  
##  AllPub:1459   Corner : 263   Gtl:1382   NAmes  :225   Norm   :1260  
##  NoSeWa:   1   CulDSac:  94   Mod:  65   CollgCr:150   Feedr  :  81  
##                FR2    :  47   Sev:  13   OldTown:113   Artery :  48  
##                FR3    :   4              Edwards:100   RRAn   :  26  
##                Inside :1052              Somerst: 86   PosN   :  19  
##                                          Gilbert: 79   RRAe   :  11  
##                                          (Other):707   (Other):  15  
##    Condition2     BldgType      HouseStyle   OverallQual    
##  Norm   :1445   1Fam  :1220   1Story :726   Min.   : 1.000  
##  Feedr  :   6   2fmCon:  31   2Story :445   1st Qu.: 5.000  
##  Artery :   2   Duplex:  52   1.5Fin :154   Median : 6.000  
##  PosN   :   2   Twnhs :  43   SLvl   : 65   Mean   : 6.099  
##  RRNn   :   2   TwnhsE: 114   SFoyer : 37   3rd Qu.: 7.000  
##  PosA   :   1                 1.5Unf : 14   Max.   :10.000  
##  (Other):   2                 (Other): 19                   
##   OverallCond      YearBuilt     YearRemodAdd    RoofStyle   
##  Min.   :1.000   Min.   :1872   Min.   :1950   Flat   :  13  
##  1st Qu.:5.000   1st Qu.:1954   1st Qu.:1967   Gable  :1141  
##  Median :5.000   Median :1973   Median :1994   Gambrel:  11  
##  Mean   :5.575   Mean   :1971   Mean   :1985   Hip    : 286  
##  3rd Qu.:6.000   3rd Qu.:2000   3rd Qu.:2004   Mansard:   7  
##  Max.   :9.000   Max.   :2010   Max.   :2010   Shed   :   2  
##                                                              
##     RoofMatl     Exterior1st   Exterior2nd    MasVnrType    MasVnrArea    
##  CompShg:1434   VinylSd:515   VinylSd:504   BrkCmn : 15   Min.   :   0.0  
##  Tar&Grv:  11   HdBoard:222   MetalSd:214   BrkFace:445   1st Qu.:   0.0  
##  WdShngl:   6   MetalSd:220   HdBoard:207   None   :864   Median :   0.0  
##  WdShake:   5   Wd Sdng:206   Wd Sdng:197   Stone  :128   Mean   : 103.7  
##  ClyTile:   1   Plywood:108   Plywood:142   NA's   :  8   3rd Qu.: 166.0  
##  Membran:   1   CemntBd: 61   CmentBd: 60                 Max.   :1600.0  
##  (Other):   2   (Other):128   (Other):136                 NA's   :8       
##  ExterQual ExterCond  Foundation  BsmtQual   BsmtCond    BsmtExposure
##  Ex: 52    Ex:   3   BrkTil:146   Ex  :121   Fa  :  45   Av  :221    
##  Fa: 14    Fa:  28   CBlock:634   Fa  : 35   Gd  :  65   Gd  :134    
##  Gd:488    Gd: 146   PConc :647   Gd  :618   Po  :   2   Mn  :114    
##  TA:906    Po:   1   Slab  : 24   TA  :649   TA  :1311   No  :953    
##            TA:1282   Stone :  6   NA's: 37   NA's:  37   NA's: 38    
##                      Wood  :  3                                      
##                                                                      
##  BsmtFinType1   BsmtFinSF1     BsmtFinType2   BsmtFinSF2     
##  ALQ :220     Min.   :   0.0   ALQ :  19    Min.   :   0.00  
##  BLQ :148     1st Qu.:   0.0   BLQ :  33    1st Qu.:   0.00  
##  GLQ :418     Median : 383.5   GLQ :  14    Median :   0.00  
##  LwQ : 74     Mean   : 443.6   LwQ :  46    Mean   :  46.55  
##  Rec :133     3rd Qu.: 712.2   Rec :  54    3rd Qu.:   0.00  
##  Unf :430     Max.   :5644.0   Unf :1256    Max.   :1474.00  
##  NA's: 37                      NA's:  38                     
##    BsmtUnfSF       TotalBsmtSF      Heating     HeatingQC CentralAir
##  Min.   :   0.0   Min.   :   0.0   Floor:   1   Ex:741    N:  95    
##  1st Qu.: 223.0   1st Qu.: 795.8   GasA :1428   Fa: 49    Y:1365    
##  Median : 477.5   Median : 991.5   GasW :  18   Gd:241              
##  Mean   : 567.2   Mean   :1057.4   Grav :   7   Po:  1              
##  3rd Qu.: 808.0   3rd Qu.:1298.2   OthW :   2   TA:428              
##  Max.   :2336.0   Max.   :6110.0   Wall :   4                       
##                                                                     
##  Electrical     X1stFlrSF      X2ndFlrSF     LowQualFinSF    
##  FuseA:  94   Min.   : 334   Min.   :   0   Min.   :  0.000  
##  FuseF:  27   1st Qu.: 882   1st Qu.:   0   1st Qu.:  0.000  
##  FuseP:   3   Median :1087   Median :   0   Median :  0.000  
##  Mix  :   1   Mean   :1163   Mean   : 347   Mean   :  5.845  
##  SBrkr:1334   3rd Qu.:1391   3rd Qu.: 728   3rd Qu.:  0.000  
##  NA's :   1   Max.   :4692   Max.   :2065   Max.   :572.000  
##                                                              
##    GrLivArea     BsmtFullBath     BsmtHalfBath        FullBath    
##  Min.   : 334   Min.   :0.0000   Min.   :0.00000   Min.   :0.000  
##  1st Qu.:1130   1st Qu.:0.0000   1st Qu.:0.00000   1st Qu.:1.000  
##  Median :1464   Median :0.0000   Median :0.00000   Median :2.000  
##  Mean   :1515   Mean   :0.4253   Mean   :0.05753   Mean   :1.565  
##  3rd Qu.:1777   3rd Qu.:1.0000   3rd Qu.:0.00000   3rd Qu.:2.000  
##  Max.   :5642   Max.   :3.0000   Max.   :2.00000   Max.   :3.000  
##                                                                   
##     HalfBath       BedroomAbvGr    KitchenAbvGr   KitchenQual
##  Min.   :0.0000   Min.   :0.000   Min.   :0.000   Ex:100     
##  1st Qu.:0.0000   1st Qu.:2.000   1st Qu.:1.000   Fa: 39     
##  Median :0.0000   Median :3.000   Median :1.000   Gd:586     
##  Mean   :0.3829   Mean   :2.866   Mean   :1.047   TA:735     
##  3rd Qu.:1.0000   3rd Qu.:3.000   3rd Qu.:1.000              
##  Max.   :2.0000   Max.   :8.000   Max.   :3.000              
##                                                              
##   TotRmsAbvGrd    Functional    Fireplaces    FireplaceQu   GarageType 
##  Min.   : 2.000   Maj1:  14   Min.   :0.000   Ex  : 24    2Types :  6  
##  1st Qu.: 5.000   Maj2:   5   1st Qu.:0.000   Fa  : 33    Attchd :870  
##  Median : 6.000   Min1:  31   Median :1.000   Gd  :380    Basment: 19  
##  Mean   : 6.518   Min2:  34   Mean   :0.613   Po  : 20    BuiltIn: 88  
##  3rd Qu.: 7.000   Mod :  15   3rd Qu.:1.000   TA  :313    CarPort:  9  
##  Max.   :14.000   Sev :   1   Max.   :3.000   NA's:690    Detchd :387  
##                   Typ :1360                               NA's   : 81  
##   GarageYrBlt   GarageFinish   GarageCars      GarageArea     GarageQual 
##  Min.   :1900   Fin :352     Min.   :0.000   Min.   :   0.0   Ex  :   3  
##  1st Qu.:1961   RFn :422     1st Qu.:1.000   1st Qu.: 334.5   Fa  :  48  
##  Median :1980   Unf :605     Median :2.000   Median : 480.0   Gd  :  14  
##  Mean   :1979   NA's: 81     Mean   :1.767   Mean   : 473.0   Po  :   3  
##  3rd Qu.:2002                3rd Qu.:2.000   3rd Qu.: 576.0   TA  :1311  
##  Max.   :2010                Max.   :4.000   Max.   :1418.0   NA's:  81  
##  NA's   :81                                                              
##  GarageCond  PavedDrive   WoodDeckSF      OpenPorchSF     EnclosedPorch   
##  Ex  :   2   N:  90     Min.   :  0.00   Min.   :  0.00   Min.   :  0.00  
##  Fa  :  35   P:  30     1st Qu.:  0.00   1st Qu.:  0.00   1st Qu.:  0.00  
##  Gd  :   9   Y:1340     Median :  0.00   Median : 25.00   Median :  0.00  
##  Po  :   7              Mean   : 94.24   Mean   : 46.66   Mean   : 21.95  
##  TA  :1326              3rd Qu.:168.00   3rd Qu.: 68.00   3rd Qu.:  0.00  
##  NA's:  81              Max.   :857.00   Max.   :547.00   Max.   :552.00  
##                                                                           
##    X3SsnPorch      ScreenPorch        PoolArea        PoolQC    
##  Min.   :  0.00   Min.   :  0.00   Min.   :  0.000   Ex  :   2  
##  1st Qu.:  0.00   1st Qu.:  0.00   1st Qu.:  0.000   Fa  :   2  
##  Median :  0.00   Median :  0.00   Median :  0.000   Gd  :   3  
##  Mean   :  3.41   Mean   : 15.06   Mean   :  2.759   NA's:1453  
##  3rd Qu.:  0.00   3rd Qu.:  0.00   3rd Qu.:  0.000              
##  Max.   :508.00   Max.   :480.00   Max.   :738.000              
##                                                                 
##    Fence      MiscFeature    MiscVal             MoSold      
##  GdPrv:  59   Gar2:   2   Min.   :    0.00   Min.   : 1.000  
##  GdWo :  54   Othr:   2   1st Qu.:    0.00   1st Qu.: 5.000  
##  MnPrv: 157   Shed:  49   Median :    0.00   Median : 6.000  
##  MnWw :  11   TenC:   1   Mean   :   43.49   Mean   : 6.322  
##  NA's :1179   NA's:1406   3rd Qu.:    0.00   3rd Qu.: 8.000  
##                           Max.   :15500.00   Max.   :12.000  
##                                                              
##      YrSold        SaleType    SaleCondition    SalePrice     
##  Min.   :2006   WD     :1267   Abnorml: 101   Min.   : 34900  
##  1st Qu.:2007   New    : 122   AdjLand:   4   1st Qu.:129975  
##  Median :2008   COD    :  43   Alloca :  12   Median :163000  
##  Mean   :2008   ConLD  :   9   Family :  20   Mean   :180921  
##  3rd Qu.:2009   ConLI  :   5   Normal :1198   3rd Qu.:214000  
##  Max.   :2010   ConLw  :   5   Partial: 125   Max.   :755000  
##                 (Other):   9






How expensive are houses?
```{r}
ggplot( homedf, aes(x = SalePrice)) +
  geom_histogram(binwidth=10000, fill = "orange", color = "grey") +
  labs(x = "Sale Price (in dollars)") +
  ggtitle("Sales Prices of Homes in Ames, IA")
```
```{r}
summary(homedf$SalePrice)
```
The sale price graph is positively skewed to the right with a good number of outliers.
The cheapest house sold for $34,900 and the most expensive for $755,000
The average sales price is $180,921, while median is $163,000

Where were the houses built?
```{r}
summary(homedf$YrSold)

```


#Finding the missing values
```{r}
sum(is.na(homedf))
# Return index of columns that have missing values
na.cols = which(colSums(is.na(homedf)) >0 )
#Break down missing values by variable
sort(colSums(sapply(homedf[na.cols], is.na)), decreasing = TRUE)

```

```{r}
ggplot(homedf, aes(x=MoSold ))+
  geom_histogram( fill ="orange" )+
  scale_x_continuous()+
  labs(x = "Month Sold")+
  ggtitle(" Monthly Sale Chart")

```

#The chart shows the home sales peak at May,June and July.

```{r}
ggplot(homedf, aes(x = GrLivArea)) +
  geom_histogram(binwidth = 100, fill = "orange",color = "grey" ) +
  labs(x = "Ground Level  sqft") +
  ggtitle(" House size ")
```

```{r}
summary(homedf$GrLivArea)
```

 The smallest house is 334 sqft and the biggest is 5642sqft
  
```{r}
ggplot(homedf, aes(x = LotArea , y = GrLivArea)) +
  geom_line() +
  labs(x = "Lot Area") +
  ggtitle(" Lot Size & Sqft ")

```

```{r}
ggplot(homedf, aes(LotArea, SalePrice)) + 
  geom_line(color="blue", size=2) +
  scale_y_continuous("Sale Price", labels = dollar) +
  scale_x_continuous("Lot area") +
  ggtitle(" Home Sale Price & Lot area") +
  theme(plot.title = element_text(size = 16, face="bold"))
```

```{r}
ggplot(homedf, aes(GrLivArea, SalePrice)) + 
  geom_line(color="blue", size=2) +
  scale_y_continuous(" Sale Price", labels = dollar) +
  scale_x_continuous("Sqft") +
  ggtitle(" Sale price and Sqft") +
  theme(plot.title = element_text(size = 16, face="bold"))
```

Both lot area and sqre footage show significant corelation with saleprice  

#Dwelling Type
The MsSubclass variable specifies the type of dwelling involved in the sale. The variable is coded as an integer and has a key that can be found in the documentation. Numbers in the 20 to 85 range represent different traditional style homes and others represent duplexes or other split homes.

```{r}
table(homedf$MSSubClass)
```

# Linear relation between Lot Frontage and Sale Price

```{r}
ggplot(homedf, aes(x = log(LotFrontage), y = SalePrice)) + 
  geom_point() +
   scale_y_continuous(labels = dollar)+ 
  geom_smooth(method = "lm")
```

```{r}
ggplot(homedf, aes(x = log(LotArea), y = SalePrice)) + 
  geom_point(alpha = 0.2) + 
  geom_smooth(method = "lm") + 
  scale_y_continuous(labels = dollar)
```


There seems to be some very high end outliers influencing the scatter plot of sale price and lot area-some lot areas are much larger than most. For this reason, we look at the natural log of lot area instead


```{r}
ggplot(homedf, aes(x = BldgType, y = SalePrice, color = OverallQual)) + geom_point()+
   scale_y_continuous(labels = dollar)+
   theme(axis.text.x = element_text(angle = 90))+
facet_wrap(~HouseStyle)
```

```{r}
ggplot(homedf, aes(x = BldgType, y = SalePrice, color = OverallCond)) + geom_point()+
   scale_y_continuous(labels = dollar)+
   theme(axis.text.x = element_text(angle = 90))+
facet_wrap(~HouseStyle)

```

Overall Quality of the home matters most.

  